---
title: "Methods and results"
author: "Jan Linnenbrink & Carles Milà"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r rmd options, echo=FALSE}
knitr::opts_chunk$set(fig.align = 'center', message = FALSE, echo=FALSE, 
                      warning = FALSE, fig.width = 8, fig.height = 3)
```

This file contains the code to reproduce figures 1, 3, 5 & 6 of the paper "kNNDM: k-fold Nearest Neighbour Distance Matching Cross-Validation for map accuracy estimation" by J Linnenbrink, C Milà, M Ludwig & H Meyer.



## Example: different clustering, NND ECDFs and W statistics

This code reproduces figure 1, where different configurations of training samples and their corresponding nearest neighbour distance (NND) empirical cumulative distribution functions (ECDF) are shown. Also, the Wasserstein statistic is shown in orange.

```{r figure 1, fig.width=10, fig.height=6, message=FALSE}
library("ggplot2")
library("sf")
library("CAST")
library("cowplot")
source("code/figures_utils.R")
source("code/sim_functions.R")

AOI <- quiet(st_read(dsn="data/species_vdata.gpkg", layer="sampling_polygon"))[1,]
AOI_part <- st_buffer(AOI, -200000)

set.seed(10)
tpoints_rand <- sim2_samples(100, "random", AOI)
set.seed(10)
tpoints_reg <- sim2_samples(100, "sregular", AOI)
set.seed(10)
tpoints_cl <-  sim2_samples(100, "sclust", AOI)
tpoints_cl$cl <- "clustered"
tpoints_rand$cl <- "random"
tpoints_reg$cl <- "regular"

tp <- rbind(tpoints_rand, tpoints_cl)

set.seed(100)
ppoints <- st_sample(AOI, 500, type = "regular") |>
  st_transform(st_crs(tpoints_rand))

set.seed(100)
k_reg <- knndm(tpoints_reg, ppoints=ppoints)
set.seed(100)
k_rand <- knndm(tpoints_rand, ppoints=ppoints)
set.seed(100)
k_clust <- knndm(tpoints_cl, k = 2,  ppoints=ppoints, maxp=0.8)

pts.size <- 1
ppts.alpha <- 0.5
base.size <- 12

Ws <- lapply(list(k_reg, k_rand,k_clust), function(x) twosamples::wass_stat(x$Gj, x$Gij))

reg_dist <- plot.knndm.integral(k_reg) +
  scale_x_continuous(labels = function(x) format(x, scientific = TRUE),n.breaks = 3, limits=c(0,4*10^5)) +
  geom_text(aes(label=paste("W=", round(Ws[[1]])), x=12^5,y=0.6), size=5, col="orange") +
  xlab("r") +
  theme_bw(base_size = base.size) +
  theme(aspect.ratio = 1, legend.position = NaN,plot.margin = unit(rep(0.2,4), "cm")) 
 
rand_dist <- plot.knndm.integral(k_rand) + 
  scale_x_continuous(labels = function(x) format(x, scientific = TRUE),n.breaks = 3, limits=c(0,4*10^5)) +
  geom_text(aes(label=paste("W=", round(Ws[[2]])), x=12^5,y=0.6), size=5, col="orange") +
  xlab("r") +
  theme_bw(base_size = base.size) +
  theme(aspect.ratio = 1, legend.position = NaN,plot.margin = unit(rep(0.2,4), "cm")) 

clust_dist <- plot.knndm.integral(k_clust) + 
  scale_x_continuous(labels = function(x) format(x, scientific = TRUE),n.breaks = 3, limits=c(0,4*10^5)) +
  xlab("r") +
  geom_text(aes(label=paste("W=", round(Ws[[3]])), x=12^5,y=0.6), size=5, col="orange") +
  theme_bw(base_size = base.size) +
  theme(aspect.ratio = 1, legend.position = NaN,plot.margin = unit(rep(0.2,4), "cm")) 
legend_bottom <- get_legend(rand_dist+
                              guides(color = guide_legend(nrow = 1)) +
                              theme(legend.position = "bottom"))

pts_cl_pl <- ggplot() +
  geom_sf(data=AOI, alpha=0) +
  geom_sf(data=ppoints, size=0.2, alpha=ppts.alpha) +
  geom_sf(data=tpoints_cl,size=pts.size) +
    scale_color_manual(values = c("blue","orange")) +
    scale_x_continuous(breaks=c(-8,0)) +
    scale_y_continuous(breaks=c(38,42)) +
  theme_bw(base_size = base.size) +
    theme(legend.position = "None",
          strip.text.x = element_text(size = 12),
          plot.margin = unit(rep(0.2,4), "cm")) +
  facet_wrap(~cl)

pts_reg_pl <- ggplot() +
  geom_sf(data=AOI, alpha=0) +
  geom_sf(data=ppoints, size=0.2, alpha=ppts.alpha) +
  geom_sf(data=tpoints_reg, size=pts.size) +
  scale_color_manual(values= rainbow(10)) +
  scale_x_continuous(breaks=c(-8,0)) +
    scale_y_continuous(breaks=c(38,42)) +
  theme_bw(base_size = base.size) +
    theme(legend.position = "None",
          strip.text.x = element_text(size = 12),
          plot.margin = unit(rep(0.2,4), "cm")) +
  facet_wrap(~cl)

pts_rand_pl <- ggplot() +
  geom_sf(data=AOI, alpha=0) +
  geom_sf(data=ppoints, size=0.2, alpha=ppts.alpha) +
  geom_sf(data=tpoints_rand, size=pts.size) +
  scale_x_continuous(breaks=c(-8,0)) +
    scale_y_continuous(breaks=c(38,42)) +
  theme_bw(base_size = base.size) +
    theme(legend.position = "None",
          strip.text.x = element_text(size = 12),
          plot.margin = unit(rep(0.2,4), "cm")) +
  facet_wrap(~cl)

top_row <- plot_grid(pts_reg_pl, pts_rand_pl, pts_cl_pl, nrow=1)
bottom_row <- plot_grid(reg_dist, rand_dist,  clust_dist, nrow=1)
comb_pl <- plot_grid(NULL, top_row, NULL, bottom_row, NULL, legend_bottom, NULL,  
                     nrow=7, rel_heights = c(-0.22,1, -0.42,1,-0.7,1,-0.45))

#ggsave("figures/1_example_ecdfs.png", comb_pl, height = 12, width=20, units="cm")
comb_pl
```


## kNNDM workflow

This code reproduces the kNNDM workflow shown in figure 3. Several numbers of clusters *q* are compared regarding their W statistic (bottom row).

```{r figure 2, fig.width=12,fig.height=6, message=FALSE}
library("ggplot2")
library("sf")
library("CAST")
library("cowplot")
source("code/figures_utils.R")

wgrid <- quiet(st_read(dsn="data/species_vdata.gpkg", layer="landscape_grid"))
AOI <- quiet(st_read(dsn="data/species_vdata.gpkg", layer="sampling_polygon")) |> 
  st_transform(st_crs(wgrid))
set.seed(10)
tpoints <-  sim2_samples(100, "sclust", AOI)
set.seed(10)
ppoints <- st_sample(AOI, 1000, type = "regular")
tpoints <- st_transform(tpoints, st_crs(ppoints))
AOI <- st_transform(AOI, st_crs(ppoints))

k <- 2
maxp <- 0.8
set.seed(10)
k_out <- k_plot(k, maxp, tpoints, ppoints)

pts_cl <- st_sf(do.call("cbind.data.frame",k_out[[1]]), geom=st_geometry(tpoints))
W <- k_out[[2]]$W

Gjstar_i <- k_out[[3]]
Gij <- k_out[[4]]
Gj <- k_out[[5]]

df <- lapply(Gjstar_i, function(x) list(x,Gij,Gj))
df <- lapply(df, function(x) {names(x) <- c("Gjstar","Gij","Gj");return(x)})

dist_plts <- mapply(function(x,y){
  plot.knndm(x) +
    ggtitle(paste0("W=", round(y))) +
    scale_x_continuous(limits=c(0,4*10^5), n.breaks = 3) +
    ylab("") +
    theme_bw() +
    theme(aspect.ratio = 1,
          legend.position = NaN,
          title=element_text(size=9))
}, x=df, y=W, SIMPLIFY = FALSE)
dst_pl <- plot_grid(plotlist = dist_plts, nrow = 1)

pts_list <- apply(st_drop_geometry(pts_cl), 2, function(x) x, simplify = FALSE)
pts_list_geo <- lapply(pts_list, function(x) st_sf(x, geom=pts_cl$geom))

pt.size <- 0.8

title_p <- k_out[[6]]

point_pl <- mapply(function(x,y) {
  ggplot() +
    ggtitle(paste0("q=", y)) +
    geom_sf(data=AOI, alpha=0) +
    geom_sf(data=x, aes(col = as.factor(x[[1]])), size=pt.size)+
    scale_color_manual(values=c("blue", "red")) +
    scale_shape_manual(values=rep(16, length(unique(x[[1]])))) +
    scale_x_continuous(breaks=c(-8,0)) +
    scale_y_continuous(breaks=c(38,42)) +
    theme_bw()+
    theme(legend.position = NaN,
          title = element_text(size=9))
}, x=pts_list_geo,y=title_p, SIMPLIFY = FALSE)

pts_pl <- plot_grid(plotlist = point_pl, nrow=1)

bottom_row <- plot_grid(dst_pl)
top_row <- plot_grid(pts_pl)
legend_bottom <- get_legend(dist_plts[[1]] +
                              guides(color = guide_legend(nrow = 1)) +
                              theme(legend.position = "bottom"))
c <- plot_grid(NULL,top_row, NULL, bottom_row, NULL, legend_bottom, nrow=6,rel_heights = c(-0.25,1,-0.35,1,-0.2,0.2))

#ggsave("figures/2_method_knndm.pdf",c, height=5, width=12)
c
```


## Simulation results

The following code reproduces figure 5, and shows the differences between cross-validated and true RMSE, R^2 and MAE for different sampling distributions.

```{r figure 3, fig.width=8, fig.height=8, message=FALSE}
library("ggplot2")
library("ggthemes")
library("cowplot")
library("dplyr")
library("tidyr")
library("forcats")

# read and process the simulation results
sim_res <- read.csv("results/sim_res.csv")

sim_diff <- data.frame(rmse_rand = sim_res$RMSE_rand - sim_res$RMSE_surf,
                       rmse_spatial = sim_res$RMSE_spatial - sim_res$RMSE_surf,
                       rmse_nndm = sim_res$RMSE_nndm - sim_res$RMSE_surf,
                       rmse_kndm = sim_res$RMSE_kndm - sim_res$RMSE_surf,
                       r2_rand =  -1*(sim_res$R2_rand - sim_res$R2_surf),
                       r2_spatial= -1*(sim_res$R2_spatial - sim_res$R2_surf),
                       r2_nndm = -1*(sim_res$R2_nndm - sim_res$R2_surf),
                       r2_kndm = -1*(sim_res$R2_kndm - sim_res$R2_surf),
                       mae_rand = sim_res$MAE_rand - sim_res$MAE_surf,
                       mae_spatial = sim_res$MAE_spatial - sim_res$MAE_surf,
                       mae_kndm = sim_res$MAE_kndm - sim_res$MAE_surf,
                       mae_nndm = sim_res$MAE_nndm - sim_res$MAE_surf,
                       dsample = sim_res$dsample) |>
  pivot_longer(!dsample) |>
  mutate(distr = case_when(dsample == "sregular" ~ "Strong\nregular",
                                     dsample == "wregular" ~ "Weak\nregular",
                                     dsample == "random" ~ "Random",
                                     dsample == "wclust" ~ "Weak\nclustered",
                                     dsample == "sclust" ~ "Strong\nclustered"),
         distr = as.factor(distr),
         distr = fct_relevel(distr, c("Strong\nregular", "Weak\nregular", "Random",
                                   "Weak\nclustered", "Strong\nclustered")),
         CV_method = case_when(grepl("rand", name) ~ "random 10-fold",
                               grepl("spatial", name) ~ "spatial 10-fold",
                               grepl("nndm", name) ~ "NNDM LOO",
                               grepl("kndm", name) ~ "kNNDM 10-fold"),
         CV_method = fct_relevel(CV_method, "random 10-fold",  "spatial 10-fold","NNDM LOO","kNNDM 10-fold"),
         group = case_when(startsWith(name, "rmse") ~ "RMSE",
                           startsWith(name, "mae") ~ "MAE",
                           startsWith(name, "r2") ~ "R^2"))

lw <- 0.6
w <- 0.45
base.size <- 15

diff_rmse <- ggplot(sim_diff[sim_diff$group=="RMSE",],
                          aes(x=distr, y=value,colour=CV_method)) +
    geom_boxplot(fill=NA, lwd=lw, width=w,position=position_dodge(0.6)) +
    scale_color_colorblind(name="CV method") +
    geom_hline(aes(yintercept=0)) +
    xlab("") +
    ylab(expression(CV - true~RMSE)) +
    theme_bw(base_size = base.size) +
    theme(legend.position = "None")

diff_R2 <- ggplot(sim_diff[sim_diff$group=="R^2", ],
                          aes(x=distr, y=value,colour=CV_method)) +
    geom_boxplot(fill=NA, lwd=lw, width=w,position=position_dodge(0.6)) +
    scale_color_colorblind(name="CV method",) +
    geom_hline(aes(yintercept=0)) +
    xlab("") +
    ylab(expression(-1* "*" * "(" * CV - true~R^2*")")) +
    theme_bw(base_size = base.size) +
    theme(legend.position = "None")

diff_MAE <- ggplot(sim_diff[sim_diff$group=="MAE",],
                          aes(x=distr, y=value,colour=CV_method)) +
    geom_boxplot(fill=NA, lwd=lw, width=w,position=position_dodge(0.6)) +
    scale_color_colorblind(name="CV method") +
    geom_hline(aes(yintercept=0)) +
    xlab("Sampling distribution") +
    ylab(expression(CV - true~MAE)) +
    theme_bw(base_size = base.size) +
    theme(legend.position = "None")

legend_right <- get_legend(diff_MAE +
                              guides(color = guide_legend(nrow = 4)) +
                              theme(legend.position = "right"))

pgr <- plot_grid(diff_rmse, NULL, NULL, NULL,  
                 diff_R2, legend_right, NULL, NULL,  
                 diff_MAE, NULL,NULL, NULL, 
                 nrow=3, rel_widths = c(1, 0.3, 0.3,-0.3), axis="l", align = "v")

ggsave("figures/3_results_sim.pdf", pgr, height=8, width=8)
pgr
```


## Association CV - True error and W statistic

The following code reproduces figure 6, which shows the association between the absolute value difference between the CV and true map accuracy statistics and W statistic.

```{r figure 4, fig.width=10,fig.height=5, message=FALSE}
library("ggplot2")
library("cowplot")
cor_res <- read.csv("results/sim_res_W.csv")
cor_diff <- data.frame(rmse_diff = cor_res$RMSE_kndm - cor_res$RMSE_surf,
                       r2_diff = cor_res$R2_kndm - cor_res$R2_surf,
                       mae_diff = cor_res$MAE_kndm - cor_res$MAE_surf,
                       ws = cor_res$WS,
                       dsample = cor_res$dsample)


m <- 0.5
b.size <- 20
lm_rmse <- lm(abs(rmse_diff)~ws, cor_diff)
r2_rmse <- round(summary(lm_rmse)$r.squared,2)
pred_rmse <- predict(lm_rmse, cor_diff)
lm_mae <- lm(abs(mae_diff)~ws, cor_diff)
r2_mae <- round(summary(lm_mae)$r.squared,2)
pred_mae <- predict(lm_mae, cor_diff)
lm_r2 <- lm(abs(r2_diff)~ws, cor_diff)
r2_r2 <- round(summary(lm_r2)$r.squared,2)
pred_r2 <- predict(lm_r2, cor_diff)

print(paste("Rsquared for Rsquared:", r2_r2, ";",
            "Rsquared for MAE:", r2_mae, ";",
            "Rsquared for RMSE:", r2_rmse))

rmse <- ggplot(data=cor_diff, aes(x=ws,y=abs(rmse_diff))) +
    geom_bin_2d(bins=30) +
    geom_line(aes(y=pred_rmse), col="black", linewidth=1.2) +
    scale_fill_viridis_b(trans="log10") +
    scale_x_continuous(n.breaks = 3,labels = function(x) format(x, scientific = FALSE)) +
    ylab(expression(abs(CV - true~RMSE))) +
    xlab("W") +
    theme_bw(base_size=b.size) +
    theme(aspect.ratio=1, legend.position = NaN,
          plot.margin = unit(unit(rep(m,4), "cm")))

mae <- ggplot(data=cor_diff, aes(x=ws,y=abs(mae_diff))) +
    geom_bin_2d(bins=30) +
    geom_line(aes(y=pred_mae), col="black", linewidth=1.2) +
    scale_fill_viridis_b(trans="log10") +
    scale_x_continuous(n.breaks = 3,labels = function(x) format(x, scientific = FALSE)) +
    ylab(expression(abs(CV - true~MAE))) +
    xlab("W") +
    theme_bw(base_size=b.size) +
    theme(aspect.ratio=1, legend.position = NaN,
          plot.margin = unit(rep(m,4), "cm"))
r2 <- ggplot(data=cor_diff, aes(x=ws,y=abs(r2_diff))) +
    geom_bin_2d(bins=30) +
    geom_line(aes(y=pred_r2), col="black", linewidth=1.2) +
    scale_fill_viridis_b(trans="log10") +
    scale_x_continuous(n.breaks = 3,labels = function(x) format(x, scientific = FALSE)) +
    ylab(expression(abs(CV - true~R^2))) +
    xlab("W") +
    theme_bw(base_size=b.size) +
    theme(aspect.ratio=1, legend.position = NaN,
        plot.margin = unit(unit(rep(m,4), "cm")))

legend_bottom <- get_legend(rmse +
                              guides(color = guide_legend(nrow = 1)) +
                              theme(legend.position = "bottom",
                                    legend.key.width  = unit(1.5, "cm"),
                                    legend.key.height  = unit(0.5, "cm"),
                                    plot.margin = unit(c(0, 0, 0, 0), "cm")))
top_row <- plot_grid(rmse, r2, mae, nrow=1, align="hv")
c <- plot_grid(top_row, NULL, legend_bottom, NULL, nrow=4,
               rel_heights = c(1,-0.35,1,-0.4), rel_widths = c(1,1,5,1))

#ggsave("figures/4_results_W_err.pdf",c, height=5, width=12)
c
```


## Computational time

The following code reproduces figure 7, which compares NNDM LOOCV and kNNDM CV regarding their computational time requirements.

```{r figure 5, fig.width=6,fig.height=3, message=FALSE}
library("ggplot2")
library("dplyr")
library("forcats")
library("tidyr")

sims <- rbind(sims_rand, sims_sclust)


time_res <- read.csv("results/time_res.csv") |> 
  mutate(distr = case_when(dsample == "sregular" ~ "Strong\nregular",
                                     dsample == "wregular" ~ "Weak\nregular",
                                     dsample == "random" ~ "Random",
                                     dsample == "wclust" ~ "Weak\nclustered",
                                     dsample == "sclust" ~ "Strong\nclustered"),
         distr = as.factor(distr),
         distr = fct_relevel(distr, c("Strong\nregular", "Weak\nregular", "Random",
                                   "Weak\nclustered", "Strong\nclustered")))

sim_stats_all <- time_res |> 
  dplyr::select(-dsample) |> 
  mutate(knndm_total = time_kndm_alg + time_kndm_mod,
         nndm_total = time_nndm_alg + time_nndm_mod) |> 
  filter(distr=="Strong\nclustered") |> 
  group_by(n_tpoints) |> 
  summarise(mean_knndm = mean(knndm_total, na.rm=TRUE),
            mean_nndm = mean(nndm_total, na.rm=TRUE),
            sd_knndm = sd(knndm_total, na.rm=TRUE),
            sd_nndm = sd(nndm_total, na.rm=TRUE),
            min_knndm = min(knndm_total, na.rm=TRUE),
            max_knndm = max(knndm_total, na.rm=TRUE),
            min_nndm = min(nndm_total, na.rm=TRUE),
            max_nndm = max(nndm_total, na.rm=TRUE))

tp <- ggplot(sim_stats_all) +
  geom_errorbar(aes(x=n_tpoints+10, ymin=mean_knndm-sd_knndm, ymax=mean_knndm+sd_knndm),
                width=10, size=0.4) +
  geom_point(aes(x=n_tpoints+10,y=mean_knndm, fill="kNNDM CV"), colour="black",
             pch=21, size=1.8) +
  geom_errorbar(aes(x=n_tpoints-10,ymin=mean_nndm-sd_nndm, ymax=mean_nndm+sd_nndm),
                width=10, size=0.4) +
  geom_point(aes(x=n_tpoints-10, y=mean_nndm, fill="NNDM LOOCV"), colour="black",
             pch=21, size=1.8) +
  scale_x_continuous(breaks=seq(100, 1000, length.out=3),
                     minor_breaks = NULL) +
  scale_y_log10() +
  ylab("time [s]") + xlab("n training points") +
  theme_bw(base_size = 16) +
  theme(legend.title = element_blank())
  
tp

ggsave("figures/5_time.pdf", tp, height=3, width=6)





```

```{r figure appendix, fig.width=8.5,fig.height=3.5, message=FALSE}

library("ggplot2")
library("dplyr")
library("forcats")
library("tidyr")


time_res <- read.csv("results/time_res.csv") |> 
  mutate(distr = case_when(dsample == "sregular" ~ "Strong\nregular",
                                     dsample == "wregular" ~ "Weak\nregular",
                                     dsample == "random" ~ "Random",
                                     dsample == "wclust" ~ "Weak\nclustered",
                                     dsample == "sclust" ~ "Strong\nclustered"),
         distr = as.factor(distr),
         distr = fct_relevel(distr, c("Strong\nregular", "Weak\nregular", "Random",
                                   "Weak\nclustered", "Strong\nclustered")))

sim_stats_extended <- time_res |> 
  dplyr::select(-dsample) |> 
  mutate(knndm_total = time_kndm_alg + time_kndm_mod,
         nndm_total = time_nndm_alg + time_nndm_mod) |> 
  group_by(n_tpoints, distr) |> 
  summarise(mean_knndm = mean(knndm_total, na.rm=TRUE),
            mean_nndm = mean(nndm_total, na.rm=TRUE),
            sd_knndm = sd(knndm_total, na.rm=TRUE),
            sd_nndm = sd(nndm_total, na.rm=TRUE),
            min_knndm = min(knndm_total, na.rm=TRUE),
            max_knndm = max(knndm_total, na.rm=TRUE),
            min_nndm = min(nndm_total, na.rm=TRUE),
            max_nndm = max(nndm_total, na.rm=TRUE))

tp_ext <- ggplot(sim_stats_extended) +
  geom_errorbar(aes(x=n_tpoints+10, ymin=mean_knndm-sd_knndm, ymax=mean_knndm+sd_knndm),
                width=10, size=0.4) +
  geom_point(aes(x=n_tpoints+10,y=mean_knndm, fill="kNNDM CV"), colour="black",
             pch=21, size=1.8) +
  geom_errorbar(aes(x=n_tpoints-10,ymin=mean_nndm-sd_nndm, ymax=mean_nndm+sd_nndm),
                width=10, size=0.4) +
  geom_point(aes(x=n_tpoints-10, y=mean_nndm, fill="NNDM LOOCV"), colour="black",
             pch=21, size=1.8) +
  scale_x_continuous(breaks=seq(100, 1000, length.out=3),
                     minor_breaks = NULL) +
  ylab("time [s]") + xlab("n training points") +
  theme_bw(base_size = 12) +
  theme(legend.title = element_blank(),
        legend.position = "bottom") +
  facet_wrap(~distr, nrow=1)
tp_ext

ggsave("figures/time_appendix.pdf", tp_ext, height=3.2, width=8.5)
```


```{r figure appendix 2, fig.width=8.5,fig.height=6, message=FALSE}

library("ggplot2")
library("dplyr")
library("forcats")
library("tidyr")


time_res <- read.csv("results/time_res_new.csv") |> 
  mutate(distr = case_when(dsample == "random" ~ "Random", dsample == "sclust" ~ "Strong\nclustered"),
         distr = as.factor(distr),
         distr = fct_relevel(distr, c("Random", "Strong\nclustered")))

t <- time_res |> 
  dplyr::select(-dsample) |> 
  mutate(knndm_total = time_kndm_alg + time_kndm_mod,
         nndm_total = time_nndm_alg + time_nndm_mod) |> 
  group_by(n_tpoints, distr) |> 
  summarise(mean_knndm_total = mean(knndm_total, na.rm=TRUE),
            mean_nndm_total = mean(nndm_total, na.rm=TRUE),
            mean_knndm_model = mean(time_kndm_mod, na.rm=TRUE),
            mean_nndm_model = mean(time_nndm_mod, na.rm=TRUE),
            mean_knndm_algorithm = mean(time_kndm_alg, na.rm=TRUE),
            mean_nndm_algorithm = mean(time_nndm_alg, na.rm=TRUE),
            standard_deviation_knndm_total = sd(knndm_total, na.rm=TRUE),
            standard_deviation_nndm_total = sd(nndm_total, na.rm=TRUE),
            standard_deviation_knndm_model = sd(time_kndm_mod, na.rm=TRUE),
            standard_deviation_nndm_model = sd(time_nndm_mod, na.rm=TRUE),
            standard_deviation_knndm_algorithm = sd(time_kndm_alg, na.rm=TRUE),
            standard_deviation_nndm_algorithm = sd(time_nndm_alg, na.rm=TRUE)) |> 
  pivot_longer(!c(n_tpoints, distr))|> 
  mutate(stat=ifelse(grepl("standard_deviation", name), "standard_deviation", "mean")) |> 
  ungroup() |> 
  tidyr::nest(.by=stat) 

names(t$data[[1]])[[4]] <- "mean"
names(t$data[[2]])[[4]] <- "sd"

td <- t$data[[1]]
td$sd <- t$data[[2]]$sd
  
sim_stats_extended <- td |>  
  mutate(CV_method=ifelse(grepl("knndm", name), "kNNDM", "NNDM"),
         what=sub(".*_", "", name)) |> 
  dplyr::select(-name)

tp_ext <- ggplot(sim_stats_extended, aes(x=n_tpoints, group=CV_method)) +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd),
                width=10, size=0.4, position = position_dodge(width=100)) +
  geom_point(aes(y=mean, fill=CV_method), colour="black", pch=21, size=1.8,
             position = position_dodge(100)) +
  scale_x_continuous(breaks=seq(100, 1000, length.out=3),
                     minor_breaks = NULL) +
  ylab("time [s]") + xlab("n training points") +
  theme_bw(base_size = 12) +
  theme(legend.title = element_blank(),
        legend.position = "bottom") +
  facet_grid(what~distr)
tp_ext

ggsave("figures/time_appendix2.pdf", tp_ext, height=6, width=8.5)

sim_stats_extended[sim_stats_extended$n_tpoints==1000 & sim_stats_extended$distr=="Strong\nclustered" & sim_stats_extended$what=="total" & sim_stats_extended$CV_method=="NNDM", "mean"][[1]] / 
sim_stats_extended[sim_stats_extended$n_tpoints==1000 & sim_stats_extended$distr=="Strong\nclustered" & sim_stats_extended$what=="total" & sim_stats_extended$CV_method=="kNNDM", "mean"][[1]]

```